Marginally-Stable Thermal Equilibria of Rayleigh-Bénard Convection

### 2021-04-16 publication_materials/ms_rbc.tex
Intro

\section{Introduction}
Rayleigh-B\'enard convection plays a foundational role in astrophysical and geophysical settings.
The resulting buoyancy-driven flows regulate heat transfer and generate large-scale vortices \cite{Couston}.
Aggressive convection, which is associated with large Rayleigh numbers $Ra$, is difficult to simulate. 
State of the art simulations performed by \cite{Zhu_2018} have reached $Ra \sim 10^{14}$ but estimates for stars are $Ra \sim 10^{20}$ \cite{Ossendrijver}. 
The scaling behavior of the Nusselt number $\Nu \sim Ra^{\beta}$ in the asymptotic ultimate regime is of particular interest.
There exists a substantial body of work pertaining to this specific topic with no general consensus \cite{Malkus_1954, Howard_1966, Kraichnan, Spiegel, Castaing, Grossman, Ahlers}.

Reduced models are particularly useful in this context because they allow us to study the problem with relatively low expense. 
Much of the published work on this topic relates to unstable exact coherent states (ECS) \cite{Waleffe, Sondak, Wen, chini_cells}. 
Simulations and analysis performed by \cite{Yalniz, Cvitanovic} suggest that chaotic simulation trajectories can be represented by a Markov chain whose long-term behavior is given by the time-weighted-average of a finite set of unstable equilibria. Should that be the case, it is crucial that we discover and classify such equilibria.

In \cite{Beaume_2015}, researchers compute ECS in parallel shear flows by deriving and solving a quasilinear formulation of the Navier-Stokes equations via multi-scale asymptotic arguments. 
This is realized by defining a mean state which, hypothetically, evolves on a slow time-scale. 
Marginal stability is required at each successive timestep to constrain the perturbation amplitudes. 
A similar strategy is employed by \cite{michel_chini_2019}, where an analytic expression for the first-order perturbations' amplitude is found by deriving a solvability condition of the second-order perturbations and integrating over the domain spatially. 
We seek to solve the Boussinesq Rayleigh–Bénard convection equations using an analogous strategy. 
The application of this procedure yields time-invariant solutions to the quasilinear problem with a range of expected and unexpected features.

@@

\section{Perturbation Evolution}
Solving the linearized system (\ref{EQ:linear1})--(\ref{EQ:linear3}) does not yield $A(t)$. 
To evolve $\bar{T}$, we assume the perturbations and mean quantities evolve according to different time-scales, as in \cite{michel_chini_2019}.
Stable modes decay away rapidly. 
Unstable modes will not persist on the slow time-scale because the advective term $\langle w'T' \rangle_x$ tends to stabilize $\bar{T}$, thereby creating a negative feedback loop. Only marginally stable modes can traverse this hypothetical time-scale separation. Therefore the amplitude function $A(t)$ must satisfy
\begin{equation}
    \max_{k_x} \{ \omega \} = 0.
\end{equation}

@@

\section{Simulations with Thermally Equilibrated Initial Conditions}
This investigation is partially motivated by the prospect of decreasing DNS times by employing MSTE initial conditions. 
Conventionally, (\ref{EQ:motion1})--(\ref{EQ:motion3}) are solved via DNS with the linear initial conditions
\begin{align}
    T(x, z)\big|_{t=0} &= 0.5 - z \nonumber \\
    \mathbf{u}(x, z)\big|_{t=0} &= \mathbf{0} \nonumber \\
    p(x, z)\big|_{t=0} &= 0 \label{EQ:linear_ic}.
\end{align}
$t$ in this section now refers to the simulation time as opposed to the thermal equilibration time parameter used previously. 
We define the MSTE initial conditions
\begin{align}
    T(x, z)\big|_{t=0} &= \bar{T}(z) + \sqrt{2} \sum_{n=1}^N  A_n \text{ Re} \Big[ \theta_n(z) e^{ik_{x_n}x} \Big] \nonumber \\
    \mathbf{u}(x, z)\big|_{t=0} &= \sqrt{2} \sum_{n=1}^N A_n \text{ Re} \Big[\Big( U_n (z) \hat{x} + W_n(z) \hat{z} \Big) e^{ik_{x_n}x} \Big] \nonumber\\
    p(x, z)\big|_{t=0} &= \sqrt{2} \sum_{n=1}^N A_n \text{ Re} \Big[P_n (z) e^{ik_{x_n}x}\Big] \label{EQ:mste_ic}
\end{align}
where $\theta_n(z), U_n(z), W_n(z), P_n(z); \, A_n; $ and $k_{x_n}$ refer to the complex eigenfunctions, amplitude, and wavenumber at the $n$th marginal mode respectively. 
The $\sqrt{2}$ factor is necessary for eigenfunction normalization.

>>>>

\section{Introduction}
Rayleigh-B\'enard convection plays a foundational role in astrophysical and geophysical settings.
The resulting buoyancy-driven flows regulate heat transfer and generate large-scale vortices \cite{Couston}.
Turbulent convection, which is associated with large Rayleigh numbers $Ra$, is difficult to simulate. 
State of the art simulations performed by \cite{Zhu_2018} have reached $Ra \sim 10^{14}$ but estimates for stars are $Ra \sim 10^{20}$ \cite{Ossendrijver}. 
\note{put in a number for earth's interior?}
The scaling behavior of the Nusselt number $\Nu \sim Ra^{\beta}$ in the asymptotic ultimate regime is of particular interest.
There exists a substantial body of work pertaining to this specific topic with no general consensus \cite{Malkus_1954, Howard_1966, Kraichnan, Spiegel, Castaing, Grossman, Ahlers}.

Absent solid evidence from direct numerical simulation, other methods have been developed to try to infer large-$Ra$ behavior or otherwise gain insight.
In the presence of other physical effects (e.g., rotation, magnetic fields), one can sometimes derive an asymptotically consistent set of reduced equations \cite{Julien2007, Julien2012}.
Reduced models are potentially useful in this context because they may allow us to study the problem with less expensive computations.
Another approach relates to unstable exact coherent states (ECS) \cite{Waleffe, Sondak, Wen, chini_cells}. 
Simulations and analysis performed by \cite{Yalniz, Cvitanovic} suggest that chaotic simulation trajectories can be represented by a Markov chain whose long-term behavior is given by the time-weighted-average over a finite set of unstable equilibria.
Should that be the case, it is crucial that we discover and classify such equilibria.

Others have turned to studying quasilinear systems.
The quasilinear approximation starts with a decomposition of all variables into a background and perturbations about this background.
The approximation is to neglect the influence of nonlinear interactions between the perturbations on the perturbations themselves \cite{marston2016}.
This renders the perturbation equations linear.
Although the quasilinear approximation greatly simplifies the problem, an additional condition must be imposed to determine the amplitude of the perturbations.
In \cite{Beaume_2015}, researchers compute ECS in parallel shear flows by deriving and solving a quasilinear formulation of the Navier-Stokes equations via multi-scale asymptotic arguments. 
They assume the background velocity evolves on a slow timescale, and to determine the perturbation amplitudes they require marginal stability at each timestep.
A similar strategy is employed by \cite{michel_chini_2019} to studying acoustic streaming.
In that work, an analytic expression for the first-order perturbation's amplitude is found by deriving a solvability condition.

In this paper we solve the Rayleigh-B\'{e}nard convection problem using an analogous strategy.
In section~\ref{sec:model} we recall the underlying equations, and in section~\ref{sec:evolution} we describe how we evolve the background temperature profile while maintaining marginal stability.
Section~\ref{sec:properties} describes the properties of the marginally-stable thermal equilibria, in particular how the Nusselt number and characteristic wavenumbers vary with the Rayleigh number.
Finally, we describe the results of simulations initialized with marginally-stable thermal equilibria in section~\ref{sec:sims}, and conclude in section~\ref{sec:Discussion}.

@@

\section{Perturbation Evolution}\label{sec:evolution}
Solving the linearized system (\ref{EQ:linear1})--(\ref{EQ:linear3}) does not yield $A(t)$. 
To evolve $\bar{T}$, we assume the perturbations and mean quantities evolve according to different time-scales, as in \cite{michel_chini_2019}.
Stable modes decay away rapidly. 
Unstable modes will not persist on the slow time-scale because the advective term $\langle w'T' \rangle_x$ tends to stabilize $\bar{T}$, thereby creating a negative feedback loop. Only marginally stable modes can traverse this hypothetical time-scale separation. Therefore the amplitude function $A(t)$ must satisfy
\begin{equation}
    \max_{k_x} \{ \omega \} = 0.
\end{equation}

@@

\section{Simulations with Thermally Equilibrated Initial Conditions}\label{sec:sims}
This investigation is partially motivated by the prospect of decreasing DNS times by employing MSTE initial conditions. 
Conventionally, (\ref{EQ:motion1})--(\ref{EQ:motion3}) are solved via DNS with the linear initial conditions
\begin{align}
    T(x, z)\big|_{t=0} &= 0.5 - z \nonumber \\
    \mathbf{u}(x, z)\big|_{t=0} &= \mathbf{0} \nonumber \\
    p(x, z)\big|_{t=0} &= 0 \label{EQ:linear_ic}.
\end{align}
$t$ in this section now refers to the simulation time as opposed to the thermal equilibration time parameter used previously. 
We define the MSTE initial conditions
\begin{align}
    T(x, z)\big|_{t=0} &= \bar{T}(z) + \sqrt{2} \sum_{n=1}^N  A_n \text{ Re} \Big[ \theta_n(z) e^{ik_{x_n}x} \Big] \nonumber \\
    \mathbf{u}(x, z)\big|_{t=0} &= \sqrt{2} \sum_{n=1}^N A_n \text{ Re} \Big[\Big( U_n (z) \hat{x} + W_n(z) \hat{z} \Big) e^{ik_{x_n}x} \Big] \nonumber\\
    p(x, z)\big|_{t=0} &= \sqrt{2} \sum_{n=1}^N A_n \text{ Re} \Big[P_n (z) e^{ik_{x_n}x}\Big] \label{EQ:mste_ic}
\end{align}
where $\theta_n(z), U_n(z), W_n(z), P_n(z); \, A_n; $ and $k_{x_n}$ refer to the complex eigenfunctions, amplitude, and wavenumber at the $n$th marginal mode respectively. 
The $\sqrt{2}$ factor is necessary for eigenfunction normalization.
